---
title: "Retail Transactions:  Monthly Report"
author: "Phoebe (Chi-Hsin) Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r seed}
set.seed(41)
```

```{r libraries}
library(data.table)
library(DT)
library(Hmisc)
```

```{r constants}

```

```{r functions}
  #function for rounding numbers
round.numerics <- function(x, digits){
  if(is.numeric(x)){
    x <- round(x = x, digits = digits)
  }
  return(x)
}

  #function for calculating linear regression coefficients
show.coef <- function(mod.lm) {
    return(data.frame(Variable = names(mod.lm$coefficients), 
        Coefficient = mod.lm$coefficients, p = summary(mod.lm)$coef[, 
            ncol(summary(mod.lm)$coef)]))
}
```

```{r load_data}
transactions <- fread(input = params$transactions.file)
customers <- fread(input = params$customers.file)
products <- fread(input = params$products.file)
views <- fread(input = params$views.file)
transactions <- fread(input = params$transactions.file)
the.year <- lubridate::year(params$ending.date)
the.month <- lubridate::month(params$ending.date)
the.day <- lubridate::day(params$ending.date)

```

```{r explore_data, eval = FALSE}

```


```{r clean_data}

```



## Part 2:  Specific Questions {.tabset}

This part of the report will be directed to product managers throughout the client's company.  The idea is to give them the useful information they need to act on the specific questions they posed.  Plan your communication accordingly.

For this report, make sure to use all of the data that was provided to you for the month.  If you do note any issues with the data (Part 3), this can be reported to the engineering team for them to resolve.

### Q1

#### Question

During the first week of the month, what were the 10 most viewed products?  Show the results in a table with the product's identifier, category, and count of the number of views.

#### Answer
```{r part2_q1}
  #merge products and views table
products_views <- merge(products, views, all.x = TRUE)
  #find views in first week of the month
week1_top10_views_products <- products_views[time < as.Date(sprintf("%s-%s-08", the.year, the.month)), .N, by = list(product_id, category)]
  #find top 10
datatable(setorderv(x = week1_top10_views_products, cols = 'N', order = -1)[1:10])


```

### Q2

#### Question

During the whole month, what were the 10 most viewed products for each category?  Show the results in separate tables by category.  Including only the product's identifier and the count of the number of views.


#### Answer {.tabset}

```{r part2_q2_overall_calculations}
products_views <- merge(views, products, all.x = TRUE)
```

##### Shirt

```{r part2_q2_shirt}
datatable(setorderv(products_views[category == 'shirt', .(total_views = .N), by = product_id], cols = 'total_views', order = -1)[1:10])
```

##### Pants

```{r part2_q2_pants}
datatable(setorderv(products_views[category == 'pants', .(total_views = .N), by = product_id], cols = 'total_views', order = -1)[1:10])
```

##### Coat

```{r part2_q2_coat}
datatable(setorderv(products_views[category == 'coat', .(total_views = .N), by = product_id], cols = 'total_views', order = -1)[1:10])
```

##### Shoes

```{r part2_q2_shoes}
datatable(setorderv(products_views[category == 'shoes', .(total_views = .N), by = product_id], cols = 'total_views', order = -1)[1:10])
```

##### Hat

```{r part2_q2_hat}
datatable(setorderv(products_views[category == 'hat', .(total_views = .N), by = product_id], cols = 'total_views', order = -1)[1:10])
```


### Q3

#### Question

What was the total revenue for each category of product during the month?  Show the results in a single table sorted in decreasing order.

#### Answer

```{r part2_q3}
  #create revenue column
transactions[, revenue := (price * quantity)]

  #merge tables
products_transactions <- merge(transactions, products, all.x = TRUE)

  #find revenues by category & order in desc. order
datatable(setorderv(products_transactions[, .('Total Revenue' = round.numerics(x = sum(revenue), digits = 3)), by = category], cols = 'Total Revenue', order = -1))

```

### Q4

#### Question

Among customers with at least one transaction, show the average, median, and standard deviation of the customers' monthly spending on the site.

#### Answer

```{r part2_q4}
  #Calculate average, median, and SD of revenue per customer
customers_spending <- transactions[, .('Avg Spend' = mean(revenue), 'Median Spend' = median(revenue), 'SD Spend' = sd(revenue))]
  
  #Round all numbers to 3 digits
datatable(customers_spending[, lapply(X = .SD, FUN = "round.numerics", digits = 3)])
```

### Q5

#### Question

What is the percentage distribution of spending by gender?  Show the amount of revenue and the percentage.

#### Answer

```{r part2_q5}
  #merge transactions and customers tables
customers_transactions <- merge(transactions, customers, all.x = TRUE)

  #Calculate revenue by gender and total revenue of all customers
gender_spend_distribution <- customers_transactions[, .('Revenue by Gender' = round.numerics(x = sum(revenue), digits = 3)), by = gender]
total_revenue <- sum(customers_transactions[, 'revenue'])

  #Calculate revenue percentage by gender
datatable(gender_spend_distribution[, 'Revenue Percentage' := round.numerics(x = `Revenue by Gender`/total_revenue, digits = 3)])


```

### Q6

#### Question

Using linear regression, what is the effect of an extra ten thousand dollars of income on monthly spending for a customer while adjusting for age, gender, and region?

#### Answer
Coefficients of the linear regression model are as shown: 
```{r part2_q6}
  #calculate and create new column for total spending per customer
customers_info_total_spend <- customers_transactions[, .(total_spending = sum(revenue), income, age, gender, region), by = customer_id]

  #conduct linear regression on total spending per customer
model <- lm(formula = total_spending ~ income + age + gender + region, data = customers_info_total_spend)

  #show linear regression model coefficients and calculate income coefficient
coef <- as.numeric(round.numerics(10000*model$coefficients['income'], 3))
datatable(as.data.frame(round.numerics(model$coefficients, 3)))

```
While adjusting for age, gender, and region, for every $10,000 increase in income, a customer's monthly spending is estimated to increase by $`r coef`.

### Q7

#### Question

Among customers who viewed at least 1 product, how many had at least one purchase during the month?  Show the total number and as a percentage of the users with a view.

#### Answer

```{r part2_q7}
  #get customer IDs that viewed a product
viewed_customer_ids <- unique(views$customer_id)

  #calculate total # and percentage of customers that viewed and purchased 
datatable(data.table(
  total_viewed_and_purchased_customers = sum(viewed_customer_ids %in% transactions$customer_id),
  viewed_customers_purchase_percentage = 
    round.numerics(sum(viewed_customer_ids %in% transactions$customer_id)/length(viewed_customer_ids), digits = 3)))
```

### Q8

#### Question

Now let's look at the viewing habits in different age groups, including 18-34, 35-49, 50-64, and 65+.  Within each group, what were the mean, median, and standard deviation for the number of unique products viewed per customer?

#### Answer

```{r part2_q8}
  #merge customers & views tables
customers_views <- merge(customers, views, by = 'customer_id', all.x = TRUE)

  #cut data table into age groups
customers_views[, age_group := cut2(x = customers_views$age, cuts = c(18, 35, 50, 64, Inf))]

  #get mean, median, sd
customers_agegroup_views_summary <- 
  customers_views[, .(age_group, length(unique(product_id))), by = customer_id][,      .(mean_unique_product_views = mean(V2),                                                       median_unique_product_views = median(V2), 
  sd_unique_product_views = sd(V2)), by = age_group]

  #order output by age group
setorderv(customers_agegroup_views_summary, cols = "age_group", order = 1)
  #round output to 3 decimal places
datatable(customers_agegroup_views_summary[, lapply(X = .SD, FUN = "round.numerics", digits = 3)])
```


### Q9 

#### Question

What is the correlation between a user's total page views and total spending?  For customers without a transaction, include their spending as zero.

#### Answer
```{r part2_q9}
    #find total spending and total page views per customer and merge
customer_spending <- transactions[, .(total_spending = sum(revenue)), by = customer_id]
customer_views <- views[, .(total_views = .N), by = customer_id]
views_customer_spending <- merge(customer_views, customer_spending, all.x = TRUE)

  #replace NA spendings with 0
views_customer_spending[is.na(views_customer_spending)] <- 0

  #calculate correlation between total page views & total spending
correlation <- round.numerics(cor(views_customer_spending$total_views, views_customer_spending$total_spending), digits = 3)
```
The correlation between total page views and total spending for the month is `r correlation`.

### Q10

#### Question

Which customer purchased the largest number of coats?  In the event of a tie, include all of the users who reached this value.  Show their identifiers and total volume.

#### Answer

```{r part2_q10}
  #calculate amount of coats bought per customer
coat_purchases_per_customer <- products_transactions[category == "coat", .(total_bought = sum(quantity)), by = customer_id]

  #get customer that bought max amount
datatable(coat_purchases_per_customer[total_bought == max(total_bought), ])
```

